Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-05-25 draft.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}
\usepackage{lineno}
\linenumbers
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

@@

%In an ideal rate measurement a volume-limited and complete sample of galaxies would be observed for a given time, and the number of SNe observed in each galaxy simply counted. In practice, apart from at very low redshift where the absolute numbers do not support robust statistical analyses, this is not possible: SNe are detected, classified and followed up based on complicated selection criteria; galaxy surveys are incomplete at faint magnitudes; the timescale of observing seasons cuts off SNe at the beginning and end of each season.
In order to measure the rate of SNe Ia per galaxy per year, we require a sample of SNe Ia, as well as a sample of all of the possible galaxies ("field" galaxies) that those SNe could have exploded in. In practice, SN and galaxy surveys do not cover the same sky area, redshift ranges, times, and have different selection biases. In this section, we introduce our SN and field galaxy samples and in Section \ref{sec:incompleteness} we describe our method of correcting the selection effects for both the SN and field galaxy samples.

@@

The results of our SED fitting are shown in Fig. \ref{fig:mass_hist_hosts_field}. The figure showcases the vastly different distributions of the two samples. SN hosts are preferentially high mass galaxies, whereas the field galaxy distribution increases down to lower masses, peaking around $10^{8.5}~\mathrm{M}_{\odot}$. The SN host stellar mass distribution is plotted twice: once as raw counts; once weighted by the host galaxy spectroscopic efficiency (Section \ref{subsec:incompletenss_SN_hosts}). As shown in Fig. \ref{fig:mass_hist_hosts_field}, the host stellar mass distribution of the DES-SN sample used in this analysis is qualitatively similar to that from the low redshift Palomar Transient Factory (PTF) sample \citep{Pan2014} as well as the large cosmological Pantheon sample \citep{Scolnic2018}. This consistency reflects that seen in the smaller spectroscopically confirmed samples presented in \citet{Wiseman2020} and \citet{Smith2020}. The field galaxy stellar mass distribution is shown on a log scale, along with the stellar mass function (SMF) from the ZFOURGE survey \citep{Tomczak2014} scaled to match at $\log (M/\mathrm{M}_{\odot}) = 10$, whose stellar masses are derived with the same photo-$z$ code, IMF and template library as used here. Above $10^{8.5}~\mathrm{M}_{\odot}$, the DES field galaxy distribution closely follows the ZFOURGE SMF for $0.5<z<0.75$, indicating that the DES sample is representative of field galaxies in this mass and redshift range.%To estimate uncertainties on the physical properties, we resample the photometry from a Gaussian distribution formed by the flux measurements and their uncertainties and rerun the SED fit. This procedure is repeated 1000 times. 
%%%% INCOMPLETENESS %%%%

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_split_sfr.png}
%    \caption{The rate per galaxy of SNe Ia as a function of stellar mass for moderately and highly star-forming galaxies, as well as passive galaxies.}%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
%    \label{fig:rate_raw_split_sfr}
%\end{figure}
%%% MODEL %%%

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_all_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$. Displayed point estimates represent the median of the posterior samples.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm}}
\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[width=.5\textwidth]{figs/fit_dtd_A_beta_tp_corner.png}
%    \caption{Posterior distributions for DTD parameters $\beta$, $A$, and $t_{\mathrm{p}}$. We display the log of $A$ for clarity. Displayed point estimates represent the median of the posterior samples.
%    %The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
%    \label{fig:corner_beta_norm_tp}}
%\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1.png}
    \caption{Rate per galaxy of SNe Ia as a function of stellar mass (green points) along with the prediction from the best fitting DTD parameters $\beta$, $A$, and $t_{\mathrm{p}}$ forward modelled through Eq. \ref{eq:galaxy_rate}. Samples from the posterior distribution of the model log rate are drawn in cyan while the black dashed line is the mean of the posterior. In red we show the simple log-linear fit from Section \ref{sec:rates}.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted}}
\end{figure}

@@

 P.W. acknowledges support from the Science and Technology Facilities Council (STFC) grant ST/R000506/1. P.W., M.Sullivan and M.Smith acknowledge support from EU/FP7-ERC grant 615929.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}
We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains, and only accept fits where $\hat{R}<1.05$.

@@

\section{Posterior distributions for DTD fits}
\label{appendix:posteriors}
In this section we present the joint and marginal posterior distributions for the various fits of the DTD presented in Section \ref{sec:split_x1_c}. Posteriors have been displayed using v0.33.0 of \texttt{ChainConsumer} \citep{Hinton2016}.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_split_c_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$ for SNe Ia split by their $c$ parameter.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm_split_c}}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fitjoint_t1_t2_corner.png}
    \caption{Posterior distributions for $t_1$ and $t_2$ of the model presented in Section \ref{subsec:split_x1}.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_t1_t2}}
\end{figure}

\input{affiliations.tex}
% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}
\usepackage{lineno}
%\linenumbers
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
%\usepackage{amssymb}	% Extra maths symbols

@@

In order to measure the rate of SNe Ia per galaxy per year, we require a sample of SNe Ia, as well as a sample of all of the possible galaxies ("field" galaxies) that those SNe could have exploded in. In practice, SN and galaxy surveys do not cover the same sky area, redshift ranges, times, and have different selection biases. In this section, we introduce our SN and field galaxy samples and in Section \ref{sec:incompleteness} we describe our method of correcting the selection effects for both the SN and field galaxy samples.

@@

The results of our SED fitting are shown in Fig. \ref{fig:mass_hist_hosts_field}. The figure showcases the vastly different distributions of the two samples. SN hosts are preferentially high mass galaxies, whereas the field galaxy distribution increases down to lower masses, peaking around $10^{8.5}~\mathrm{M}_{\odot}$. The SN host stellar mass distribution is plotted twice: once as raw counts; once weighted by the host galaxy spectroscopic efficiency (Section \ref{subsec:incompletenss_SN_hosts}). As shown in Fig. \ref{fig:mass_hist_hosts_field}, the host stellar mass distribution of the DES-SN sample used in this analysis is qualitatively similar to that from the low redshift Palomar Transient Factory (PTF) sample \citep{Pan2014} as well as the large cosmological Pantheon sample \citep{Scolnic2018}. This consistency reflects that seen in the smaller spectroscopically confirmed samples presented in \citet{Wiseman2020} and \citet{Smith2020}. The field galaxy stellar mass distribution is shown on a log scale, along with the stellar mass function (SMF) from the ZFOURGE survey \citep{Tomczak2014} scaled to match at $\log (M/\mathrm{M}_{\odot}) = 10$, whose stellar masses are derived with the same photo-$z$ code, IMF and template library as used here. Above $10^{8.5}~\mathrm{M}_{\odot}$, the DES field galaxy distribution closely follows the ZFOURGE SMF for $0.5<z<0.75$, indicating that the DES sample is representative of field galaxies in this mass and redshift range.

@@

%%% MODEL %%%

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_all_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$. Displayed point estimates represent the median of the posterior samples.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm}}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1.png}
    \caption{Rate per galaxy of SNe Ia as a function of stellar mass (green points) along with the prediction from the best fitting DTD parameters $\beta$, $A$, and $t_{\mathrm{p}}$ forward modelled through Eq. \ref{eq:galaxy_rate}. Samples from the posterior distribution of the model log rate are drawn in cyan while the black dashed line is the mean of the posterior. In red we show the simple log-linear fit from Section \ref{sec:rates}.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:rate_fitted}}
\end{figure}

@@

 P.W. acknowledges support from the Science and Technology Facilities Council (STFC) grant ST/R000506/1. M.Sullivan and M.Smith acknowledge support from EU/FP7-ERC grant 615929.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/fit_fine_BC03_corner.png}
    \caption{Joint posterior distribution for the slope and intercept of the linear fit to the SN rate per galaxy per year as a function of stellar mass (Section \ref{sec:rates}; Fig. \ref{fig:rate_fitted}).}
    \label{fig:corner_slope_int}
\end{figure}

We model the relation with the linear relationship: 
\begin{equation}
    R_{\mathrm{G}} = \frac{\mathrm{d}R}{\mathrm{d}M_*} M_* + c \,,
\label{eq:rate_fit}
\end{equation}
where $dR/dM_*$ signifies the change of the rate of SNe as a function of the stellar mass, $c$ is a constant that sets the normalisation of the rate. We fit the model assuming a normal likelihood, such that the observed rate in each stellar mass bin is itself modelled as a Gaussian distribution described by the mean and standard deviation of the data in that bin. We adopt weakly informative normal priors on the slope and intercept: $p\left(\frac{\mathrm{d}R}{\mathrm{d}M_*}\right) \sim \mathcal{N}\left(0,5\right)$ and  $p\left(c\right) \sim \mathcal{N}\left(-12,5\right)$ respectively. We sample using 4 chains, each with 2000 warm up and 2000 sampling iterations. We report parameter estimates based on the mean and standard deviation of their posterior samples. We use the $\hat{R}$ diagnostic of \citep{Vehtari2019} to assess the convergence of MC chains, and only accept fits where $\hat{R}<1.05$.

@@

\section{Posterior distributions for DTD fits}
\label{appendix:posteriors}

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fit_dtd_split_c_A_beta_corner.png}
    \caption{Posterior distributions for DTD slope $\beta$ and normalisation $A$ for SNe Ia split by their $c$ parameter.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_beta_norm_split_c}}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/fitjoint_t1_t2_corner.png}
    \caption{Posterior distributions for $t_1$ and $t_2$ of the model presented in Section \ref{subsec:split_x1}.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citetalias{Childress2014} model}
    \label{fig:corner_t1_t2}}
\end{figure}
In this section we present the joint and marginal posterior distributions for the various fits of the DTD presented in Section \ref{sec:split_x1_c}. Posteriors have been displayed using v0.33.0 of \texttt{ChainConsumer} \citep{Hinton2016}.
\clearpage
\onecolumn
\parbox{\textwidth}{
\scriptsize
$^{1}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK\\
$^{2}$ Universit\'e de Lyon 1, CNRS/IN2P3, Institut de Physique des Deux Infinis, 69622 Villeurbanne Cedex, France\\
$^{3}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK\\
$^{4}$ Department of Physics, Duke University Durham, NC 27708, USA\\
$^{5}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia\\
$^{6}$ NASA Einstein Fellow\\
$^{7}$ Center for Astrophysics $\vert$ Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{8}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia\\
$^{9}$ PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{10}$ Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA\\
$^{11}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA\\
$^{12}$ Centre for Gravitational Astrophysics, College of Science, The Australian National University, ACT 2601, Australia\\
$^{13}$ Universit\'e Clermont Auvergne, CNRS/IN2P3, LPC, F-63000 Clermont-Ferrand, France\\
$^{14}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain\\
$^{15}$ INAF, Osservatorio Astronomico di Trieste, I-34143 Trieste, Italy\\
$^{16}$ Centre for Astrophysics \& Supercomputing, Swinburne University of Technology, Victoria 3122, Australia\\
$^{17}$ Sydney Institute for Astronomy, School of Physics, A28, The University of Sydney, NSW 2006, Australia\\
$^{18}$ Cerro Tololo Inter-American Observatory, NSF's National Optical-Infrared Astronomy Research Laboratory, Casilla 603, La Serena, Chile\\
$^{19}$ Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil\\
$^{20}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA\\
$^{21}$ Instituto de F\'{i}sica Te\'orica, Universidade Estadual Paulista, S\~ao Paulo, Brazil\\
$^{22}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France\\
$^{23}$ Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France\\
$^{24}$ Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK\\
$^{25}$ Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA\\
$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA\\
$^{27}$ Instituto de Astrofisica de Canarias, E-38205 La Laguna, Tenerife, Spain\\
$^{28}$ Universidad de La Laguna, Dpto. Astrofísica, E-38206 La Laguna, Tenerife, Spain\\
$^{29}$ Center for Astrophysical Surveys, National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA\\
$^{30}$ Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA\\
$^{31}$ Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain\\
$^{32}$ Astronomy Unit, Department of Physics, University of Trieste, via Tiepolo 11, I-34131 Trieste, Italy\\
$^{33}$ INAF-Osservatorio Astronomico di Trieste, via G. B. Tiepolo 11, I-34143 Trieste, Italy\\
$^{34}$ Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy\\
$^{35}$ Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil\\
$^{36}$ Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA\\
$^{37}$ Department of Physics, IIT Hyderabad, Kandi, Telangana 502285, India\\
$^{38}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
$^{39}$ Institute of Theoretical Astrophysics, University of Oslo. P.O. Box 1029 Blindern, NO-0315 Oslo, Norway\\
$^{40}$ Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain\\
$^{41}$ Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain\\
$^{42}$ Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain\\
$^{43}$ Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{44}$ Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{45}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
$^{46}$ Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA\\
$^{47}$ Department of Physics, The Ohio State University, Columbus, OH 43210, USA\\
$^{48}$ Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 Munich, Germany\\
$^{49}$ Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85748 Garching, Germany\\
$^{50}$ Department of Astronomy/Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA\\
$^{51}$ Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia\\
$^{52}$ Lowell Observatory, 1400 Mars Hill Rd, Flagstaff, AZ 86001, USA\\
$^{53}$ George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA\\
$^{54}$ Department of Astronomy, The Ohio State University, Columbus, OH 43210, USA\\
$^{55}$ Radcliffe Institute for Advanced Study, Harvard University, Cambridge, MA 02138\\
$^{56}$ Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain\\
$^{57}$ Physics Department, 2320 Chamberlin Hall, University of Wisconsin-Madison, 1150 University Avenue Madison, WI  53706-1390\\
$^{58}$ Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA\\
$^{59}$ Department of Physics and Astronomy, Pevensey Building, University of Sussex, Brighton, BN1 9QH, UK\\
$^{60}$ Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831\\
$^{61}$ Universit\"ats-Sternwarte, Fakult\"at f\"ur Physik, Ludwig-Maximilians Universit\"at M\"unchen, Scheinerstr. 1, 81679 M\"unchen, Germany\\
}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}
